---
title: "China's Role in UN Sanctions: Reassessing China's Adherence to the Liberal International Order with CEC Data"
author: "David Cardaronella"
date: "2026-03-02"
format: html
editor: visual
embed-resources: true
theme: minty
---

## Voting

#### Packages Used

```{r}
#| message: false
#| warning: false
#| code-fold: true

library(data.table)
library(ggraph)
library(gt)
library(igraph)
library(patchwork)
library(pscl)
library(readxl)
library(tidyverse)
```


#### Reading in Resolutions

```{r reading in resolutions}
#| code-fold: true

resolutions <- 
  fread("CR-UNSC_2024-05-19_ALL_CSV_FULL.csv") |>
  as_tibble() |>
  filter(year %in% c(1972:2024)) |> 
  select(
    year, date, res_no, iso_name, m49_region_sub, topic, subjects, chapter7,
    vote_summary, vote_detail, vote_yes, vote_no, vote_abstention, vote_nonvote
  ) |> 
  rename("names" = iso_name, "subregion" = m49_region_sub)
  
resolutions

```

#### Adding Voting Records of P5 to Resolutions

```{r adding voting records of p5 to resolutions}
#| code-fold: true

resolutions_votes <-
  resolutions |>
  mutate(
    china_yes = case_when(
      str_detect(vote_detail, "Y CHINA") ~ 1,
      TRUE ~ 0
    ),
    china_no = case_when(
      str_detect(vote_detail, "N CHINA") ~ 1,
      TRUE ~ 0
    ),
    china_ab = case_when(
      str_detect(vote_detail, "A CHINA") ~ 1,
      TRUE ~ 0
    ),
    france_yes = case_when(
      str_detect(vote_detail, "Y FRANCE") ~ 1,
      TRUE ~ 0
    ),
    france_no = case_when(
      str_detect(vote_detail, "N FRANCE") ~ 1,
      TRUE ~ 0
    ),
    france_ab = case_when(
      str_detect(vote_detail, "A FRANCE") ~ 1,
      TRUE ~ 0
    ),
    us_yes = case_when(
      str_detect(vote_detail, "Y UNITED STATES") ~ 1,
      TRUE ~ 0
    ),
    us_no = case_when(
      str_detect(vote_detail, "N UNITED STATES") ~ 1,
      TRUE ~ 0
    ),
    us_ab = case_when(
      str_detect(vote_detail, "A UNITED STATES") ~ 1,
      TRUE ~ 0
    ),
    uk_yes = case_when(
      str_detect(vote_detail, "Y UNITED KINGDOM") ~ 1,
      TRUE ~ 0
    ),
    uk_no = case_when(
      str_detect(vote_detail, "N UNITED KINGDOM") ~ 1,
      TRUE ~ 0
    ),
    uk_ab = case_when(
      str_detect(vote_detail, "A UNITED KINGDOM") ~ 1,
      TRUE ~ 0
    ),
    russia_yes = case_when(
      str_detect(vote_detail, "Y USSR") ~ 1,
      str_detect(vote_detail, "Y RUSSIAN FEDERATION") ~ 1,
      TRUE ~ 0
    ),
    russia_no = case_when(
      str_detect(vote_detail, "N USSR") ~ 1,
      str_detect(vote_detail, "N RUSSIAN FEDERATION") ~ 1,
      TRUE ~ 0
    ),
    russia_ab = case_when(
      str_detect(vote_detail, "A USSR") ~ 1,
      str_detect(vote_detail, "A RUSSIAN FEDERATION") ~ 1,
      TRUE ~ 0
    )
  ) |>
  select(
    year, date, res_no, names, subregion, topic, subjects, chapter7, china_yes,
    china_no, china_ab, france_yes, france_no, france_ab, us_yes, us_no, us_ab,
    uk_yes, uk_no, uk_ab, russia_yes, russia_no, russia_ab, vote_yes, vote_no,
    vote_abstention, vote_nonvote
  )

resolutions_votes

# Usually a vote of "no" by a P5 member vetoes and so prevents a resolution, but when a resolution is over a non-substantive matter, such as calling an Emergency Special Session of the General Assembly, it cannot be vetoed. 

# Russia has 2 no votes in here, both against calling an Emergency Special Session. Those are the only no votes.

```

#### All passed and Vetoed Sanctions

```{r All passed and vetoed sanctions}
#| code-fold: true

# All passed resolutions over sanctions

sanctions <- 
  resolutions_votes |> 
  filter(
    str_detect(subjects, regex("SANCTIONS")),
    chapter7 == TRUE
  )


# All vetoed sanctions

vetoed <-
  fread("DPPA-SCVETOES.csv") |> 
  as_tibble() |>
  rename(
    "china_no" = china,
    "france_no" = france,
    "russia_no" = russian_federation_ussr,
    "us_no" = united_states,
    "uk_no" = united_kingdom
  ) |> 
  select(
    year, date, "draft_res#", short_agenda, china_no, france_no, russia_no,
    us_no, uk_no
  ) |>
  rename("draft_no" = "draft_res#", "topic" = "short_agenda") |> 
  filter(year %in% c(1972:2024)) |>
  mutate(success = 0) |> 
  arrange(year) |> 
  slice(-5) # row 5 vetoed by US, not China and Russia, should be removed

# If a state did not veto, they either voted yes or abstained, no non-votes

# Success is 0 for vetoes, because the sanction threat contained within the draft could never be imposed, the threat failed.


# Joining passed drafts (resolutions) and vetoed drafts

sanctions_votes <- 
  full_join(
    sanctions, 
    vetoed, 
    by = join_by(year, date, topic, china_no, france_no, russia_no, us_no, uk_no)
  ) |> 
  select(
    year, date, res_no, draft_no, names, topic, subjects, china_yes, china_no,
    china_ab, france_yes, france_no, france_ab, us_yes, us_no, us_ab, uk_yes,
    uk_no, uk_ab, russia_yes, russia_no, russia_ab
  ) |> 
  arrange(date)

sanctions_votes

```

## Cases

#### CMPS Case Data on UNSC Resolutions

```{r CMPS UNSC Case Data}
#| code-fold: true

CMPS <-
  read_csv("CESv1 Sender CMPS.csv") |>
  filter(UNSC == 1) |> 
  mutate(
    date = as.IDate(
      str_glue("{startyear}-{sprintf('%02d', startmonth)}-{sprintf('%02d', startday)}")
    )
  ) |>
  select(
    `Case ID`, caseid_name, date, ccode_target, state2, finaloutcome, success,
    `Narrative Link`
  ) |> 
  rename(
    "caseid" = `Case ID`, 
    "link" = `Narrative Link`, 
    "target" = ccode_target,
    "target_name" = state2
  ) |> 
  mutate(cmps = 1)

CMPS

```

#### EUSANCT Case Data on UNSC Resolutions

```{r EUSANCT UNSC Case Data}
#| code-fold: true

EUSANCT <- 
  read_xls("EUSANCT_Dataset_Case-level.xls") |> 
  filter(str_detect(threatUN_id, "UNSC")) |> 
  select(
    caseid, startdateUN, targetstate, targetstate_name, finaloutcome, success, link
  ) |>
  rename(
    "date" = "startdateUN", 
    "target" = "targetstate", 
    "target_name" = "targetstate_name"
  ) |>
  mutate(
    date = as.IDate(date, format = "%d%b%Y"),
    caseid = as.numeric(caseid),
    eusanct = 1
  )

EUSANCT

```

#### Joining CMPS and EUSANCT Case Data

```{r CMPS and EUSANCT UNSC Case Data}
#| code-fold: true

cases <- 
  full_join(
    EUSANCT, 
    CMPS, 
    by = join_by("caseid", "date", "target", "target_name", 
                 "finaloutcome", "success", "link")
  ) |> 
  select(caseid, caseid_name, date, target, target_name,
         finaloutcome, success, link, eusanct, cmps) |>
  mutate(
    cmps = replace_na(cmps, 0),
    eusanct = replace_na(eusanct, 0)
  ) |>
  arrange(date) |> 
  slice(-1) |> # All other rows besides first fit within 1972-2024
  mutate(row = row_number())  # To identify original row number for later
  
cases

```

#### Fixing Case Data

```{r cases fixed}
#| code-fold: true

# To join resolutions_votes with cases, cases needs to have a res_no variable to match with. We cannot match by date, because some rows have the same date and go to different resolutions.

# Row 2 date needs to be 1992-05-30

# Row 6 date needs to be 1992-11-30

# Row 10 date needs to be 1994-04-21

# Row 11 had no votes, should be removed

# Row 21 is same resolution as Row 20

# Row 24-30 are duplicates and are not resolutions, should be removed

# Row 32 is same resolution as Row 31

# Row 34 is same resolution as 33

# Row 36 is same resolution as Row 37

# Row 38 is not a resolution, should be removed

# Row 39 date needs to be 2001-09-28

# Row 51 is same resolution as Row 53

# Row 55 is same resolution as Row 54

# Row 56 date needs to be 2006-06-20

# Row 67 date needs to be 2011-03-30

# Row 70 is same resolution as Row 69

# Row 74 is same resolution as Row 73

# Row 75 is same resolution as Row 76 and coded wrong, should be removed

# Row 77 counts the threat of sanction as a success, and since the results followed the imposition of the sanction Res 2085, although the imposition was not mentioned, I will count Res 2085 as a success. The date needs to be 2012-12-20

# Row 81 date needs to be 2015-03-03

# Row 83 has no resolution to back it up, should be removed

rows_to_remove <- c(11, 24, 25, 26, 27, 28, 29, 30, 38, 75, 83)

cases_fixed <- 
  cases[-rows_to_remove, ] |> 
  mutate(
    date = case_when(
      row == 2  ~ as.IDate("1992-05-30"),
      row == 6  ~ as.IDate("1992-11-30"),
      row == 10 ~ as.IDate("1994-04-21"),
      row == 39 ~ as.IDate("2001-09-28"),
      row == 56 ~ as.IDate("2006-06-20"),
      row == 67 ~ as.IDate("2011-03-30"),
      row == 77 ~ as.IDate("2012-12-20"),
      row == 81 ~ as.IDate("2015-03-03"),
      TRUE ~ date
    )
  )

cases_fixed <- 
  cases_fixed |> 
  left_join(
    tibble::tibble(
      row = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21,
              22, 23, 31, 32, 33, 34, 35, 36, 37, 39, 40, 41, 42, 43, 44, 45, 46, 47,
              48, 49, 50, 51, 52, 53, 54, 55, 56, 57, 58, 59, 60, 61, 62, 63, 64, 65,
              66, 67, 68, 69, 70, 71, 72, 73, 74, 76, 77, 78, 79, 80, 81, 82),
      res_no = c(661, 757, 713, 733, 788, 792, 861, 864, 873, 912, 1011, 1072,
                 1127, 1132, 1134, 1156, 1173, 1221, 1227, 1227, 1267, 1270, 1298,
                 1298, 1304, 1304, 1306, 1343, 1343, 1373, 1389, 1407, 1464, 1478,
                 1483, 1493, 1521, 1546, 1556, 1572, 1584, 1636, 1596, 1636, 1640,
                 1640, 1689, 1695, 1697, 1737, 1747, 1803, 1807, 1844, 1925, 1929,
                 1962, 1975, 1980, 1989, 1989, 2009, 2036, 2046, 2046, 2048, 2085,
                 2094, 2121, 2127, 2206, 2201)
    ),
    by = "row"
  ) |>
  slice(-2)  # remove row with no success rating

cases_fixed

```

## Case-Level Voting

#### Case Data Joined with Voting Data

```{r CMPS and EUSANCT UNSC Case Data with voting data}
#| code-fold: true

# cases_fixed is joined with resolutions_votes, only resolutions over sanctions (invoking chapter VII) are included

cases_votes <-
  left_join(cases_fixed, resolutions_votes, by = join_by("res_no", "date")) |>
  # Joining by unique res_no-date combination
  filter(chapter7 == TRUE) |> 
  select(
    year, caseid, caseid_name, res_no, date, chapter7, target, target_name, topic,
    china_yes, china_no, china_ab, france_yes, france_no, france_ab, us_yes, us_no,
    us_ab, uk_yes, uk_no, uk_ab, russia_yes, russia_no, russia_ab, vote_yes, vote_no,
    vote_abstention, vote_nonvote, finaloutcome, success, link, cmps, eusanct
  ) |>
  arrange(date)

cases_votes

```

#### Deciding To Not Include TIES Dataset

```{r Deciding To Not Include TIES Dataset}
#| code-fold: true

# Should we include data from TIES in addition to data from CMPS and EUSANCT?

cases_votes # Combines CMPS and EUSANCT with voting data
# 2 abstains from china included here, only 1 more when not filtering for resolutions invoking Chapter VII

resolutions_votes |> filter(china_ab == 1)

# no vetoes contained in EUSANCT, CMPS. The significance of China's non-support came from DPPA-SCVETOES for vetoes, combined with case data from EUSANCT and CMPS to look at other successes or failures where China voted yes (51) or abstained (2).

TIES <- 
  read_excel("TIESv4-1.xls") |> 
  as_tibble() |>
  filter(startyear %in% c(1972:2005)) |> 
  mutate(
    date = as.IDate(
      str_glue("{startyear}-{sprintf('%02d', startmonth)}-{sprintf('%02d', startday)}"))
  ) |> 
  drop_na(date) |> 
  select(date, primarysender, targetstate)

TIES

inner_join(vetoed, TIES, by = "date")
# 2 vetoes contained in TIES, neither over sanctions, so no value in including it

```

## Incomplete Correlation

#### Overview of How Often China Votes Yes

```{r Case-level Sanction Success}
#| code-fold: true

# Out of 2415 resolutions, 2196 were ones China voted yes on: 91%
resolutions_votes |> summarize(percent_china_yes = sum(china_yes) / n())

# Out of 364 non-vetoed sanctions, 318 were ones China voted yes on: 87%
sanctions |> summarize(percent_china_yes = sum(china_yes) / n())

# Out of 57 sanctions with success data, 55 were ones China voted yes on: 96%
# The other 2 were abstains
cases_votes |> summarize(cases_votes = sum(china_yes) / n())

# Out of 35 successful sanctions, 34 were ones China voted yes on: 97%
cases_votes |> 
  filter(success == 1) |> 
  summarize(percent_china_yes = sum(china_yes) / n())

# Out of 22 unsuccessful sanctions, 21 were ones China voted yes on: 95%
cases_votes |> 
  filter(success == 0) |> 
  summarize(percent_china_yes = sum(china_yes) / n())

```

#### Without Vetoes, We Fail to Find Influence of China's Voting

```{r Unsuccessful Logistic Regression without vetoes}
#| code-fold: true

#p-value = 0.74
log_reg0 <- glm(success ~ china_ab, data = cases_votes, family = binomial)

#log_reg1 <- glm(success ~ china_yes, data = cases_votes, family = binomial)
#log_reg2 <- glm(success ~ france_yes, data = cases_votes, family = binomial)
#log_reg3 <- glm(success ~ france_ab, data = cases_votes, family = binomial)
#log_reg4 <- glm(success ~ us_yes, data = cases_votes, family = binomial)
#log_reg5 <- glm(success ~ us_ab, data = cases_votes, family = binomial)
#log_reg6 <- glm(success ~ uk_yes, data = cases_votes, family = binomial)
#log_reg7 <- glm(success ~ uk_ab, data = cases_votes, family = binomial)
#log_reg8 <- glm(success ~ russia_yes, data = cases_votes, family = binomial)
#log_reg9 <- glm(success ~ russia_ab, data = cases_votes, family = binomial)

summary(log_reg0)
#summary(log_reg1)
#summary(log_reg2)
#summary(log_reg3)
#summary(log_reg4)
#summary(log_reg5)
#summary(log_reg6)
#summary(log_reg7)
#summary(log_reg8)
#summary(log_reg9)

# Running this test with China, Russia, France, the UK, and the US, none are statistically significant

# The issue is that there are only 2 abstains from China, 1 abstain from France, 1 abstain from Russia, and 0 abstains from the US and the UK. The only resolutions which get passed are those which do not have any vetoes, so there cannot be any "no" votes except for non-substantive matters. All other votes are yes votes.

cases_votes |> 
  summarize(
    china_ab = sum(china_ab),
    france_ab = sum(france_ab),
    us_ab = sum(us_ab),
    uk_ab = sum(uk_ab),
    russia_ab = sum(russia_ab)
  )

# To fully analyze the effect of China's non-support on UNSC sanction success, we need to include China's vetoes, not just its abstains

```

## Correlation with Vetoes

#### Case-Level Data with China's Vetoes

```{r Case level data with China vetoes}
#| code-fold: true

# Total Vetoes for each of the P5

vetoed |> 
  summarize(
    china_veto = sum(china_no),
    france_veto = sum(france_no),
    russia_veto = sum(russia_no),
    us_veto = sum(us_no),
    uk_veto = sum(uk_no)
  )


# Analyzing which of China's 19 vetoes were over sanctions

#1: Not a sanction (Not allowing Bangladesh to join UN)
#2: Not a sanction (Blocked peacekeeping mission to Guatemala over Taiwan recognition)
#3: Not a sanction (Blocked peacekeeping mission to Yugoslav over Taiwan recognition)
#4: Not a sanction
#5: Sanction against Zimbabwe
#6: Not a sanction against Syria (No Chapter VII)
#7: Sanction against Syria
#8: Not a sanction against Syria (No Chapter VII)
#9: Sanction against Syria
#10: Not a sanction against Syria (No Chapter VII)
#11: Sanction against Syria	
#12: Not a sanction against Syria (No Chapter VII)
#13: Not a sanction against Syria (No Chapter VII)
#14: Not a sanction against Venezuela (No Chapter VII)
#15: Not a sanction against Syria (No Chapter VII)
#16: Not a sanction against Syria (No Chapter VII)
#17: Sanction against North Korea
#18: Not a sanction against Hamas (No Chapter VII)
#19: Not a sanction against Israel, Hamas, and Houthis (No Chapter VII)

# 5 of China's vetoes were over resolutions which, if passed, would have been sanctions

china_vetoes <- 
  vetoed |> 
  filter(china_no == 1) |> 
  slice(c(5, 7, 9, 11, 17))

# The vetoed sanctions which were not vetoed by China, which China abstained on, were not analyzed. Though China did not support these sanctions, if it did not veto them, the sanctions failures were caused solely by whichever state did veto the sanction. The sanctions failures were not caused by China's abstaining, and so analyzing the effect of China's non-support in those cases would be misleading


# Joining Case-level data with China's veto data (vetoed sanctions are by definition unsuccessful)

china_votes_vetoes <- 
  full_join(
    cases_votes, 
    china_vetoes, 
    by = join_by(year, date, topic, china_no, france_no, 
                 russia_no, us_no, uk_no, success)
  ) |> 
  mutate(
    china_ns = if_else(china_no == 1 | china_ab == 1, 1, 0),
    china_yes = replace_na(china_yes, 0),
  ) |> 
  select(year, date, topic, china_yes, china_ns, success) |> 
  arrange(year)

china_votes_vetoes

# china_ns represents China's non-support for a sanction. It is a combination of abstains and vetoes, "1" if China abstained or vetoed, and "0" if China voted yes.

```

#### With Vetoes, We Find Influence of China's Voting

```{r Successful Logistic Regression including vetoes}
#| code-fold: true

# Logistic Regression of the effect of China's non-support for a sanction (no or abstain) on success

log_reg_vv <- glm(success ~ china_ns, data = china_votes_vetoes, family = binomial)

summary(log_reg_vv)

# Statistically-significant at the 5% level (p-value: 0.04)

# Based on this, the model would be: logit(P(Success = 1)) = 0.48 − 2.27ChinaNS

# Predicted probability of success when China supports a sanction is 0.618

1 / (1 + exp(-(log_reg_vv$coefficients["(Intercept)"])))

# Predicted probability of success when China does not support a sanction is 0.143

1 / (1 + exp(-(log_reg_vv$coefficients["(Intercept)"] +
                 log_reg_vv$coefficients["china_ns"])))

```

#### How Much Variation in Success is Explained?

```{r Tjur R-squared}
#| code-fold: true

pred_probs <- predict(log_reg_vv, type = "response")

mean_success <- mean(pred_probs[china_votes_vetoes$success == 1])
mean_failure <- mean(pred_probs[china_votes_vetoes$success == 0])

tjur_r2 <- mean_success - mean_failure
tjur_r2

# Around 9% of variation in sanction success explained by China's non-support

```

## Vote similarity

#### Vote Similarity Function

```{r Vote Similarity function}
#| code-fold: true

# Custom function for vote similarity, works with resolutions_votes and subsets

vote_similarity <- function(data) {
  
  sim_data <- function(data, state1, state2) {
    
    yes1 <- sym(str_c(state1, "_yes"))
    no1  <- sym(str_c(state1, "_no"))
    ab1  <- sym(str_c(state1, "_ab"))
    
    yes2 <- sym(str_c(state2, "_yes"))
    no2  <- sym(str_c(state2, "_no"))
    ab2  <- sym(str_c(state2, "_ab"))
    
    data |> 
      transmute(
        # agreements: yes-yes, no-no, no-ab, ab-no, ab-ab
        agreement = ((!!yes1 & !!yes2) |
                       (!!no1 & !!no2) |
                      ((!!no1 & !!ab2) |
                      (!!ab1 & !!no2)) |
                       (!!ab1 & !!ab2)),
        # disagreements: yes-no, yes-ab, no-yes, ab-yes
        disagreement = ((!!yes1 & !!no2) |
                        (!!yes1 & !!ab2) |
                        (!!no1 & !!yes2) |
                        (!!ab1 & !!yes2)),
        # mutual votes: both voting
        mutual_vote = ((!!yes1 | !!no1 | !!ab1) & (!!yes2 | !!no2 | !!ab2))
      ) |> 
      summarize(
        agreements = sum(agreement, na.rm = TRUE),
        disagreements = sum(disagreement, na.rm = TRUE),
        mutual_votes = sum(mutual_vote, na.rm = TRUE),
        similarity = ((agreements - disagreements) / mutual_votes + 1) / 2,
      ) |> 
      mutate(pair = str_c(str_to_upper(state1), "-", str_to_upper(state2))) |>
      select(pair, similarity)
    
    }

  states <- c("china", "us", "russia", "france", "uk")
  all_pairs <- combn(states, 2, simplify = FALSE)

  vote_sim <- 
    map_dfr(all_pairs, ~ sim_data(data, .x[1], .x[2])) |> 
    arrange(similarity) |>
    select(pair, similarity)
  
  print(vote_sim)
}

general_similarity <- 
  vote_similarity(sanctions_votes)

# agreements and disagreements = mutual votes

# the number of mutual votes across all of the P5 is not equal to the total number of votes, because of mutual non-votes (both not voting), and non-mutual votes (one state not voting)

# The only disagreements in resolutions_votes are those for yes-ab or ab-yes, besides the two no votes Russia has cast over calling an Emergency Special Session. However, sanctions_votes includes vetoed, so it adds disagreements for yes-no and no-yes

```

#### Vote Similarity by LIO Subject

```{r Vote Similarity by LIO Subject}
#| code-fold: true

# Subjects representing elements of the Liberal International Order:

# Security

law <- 
  sanctions_votes |>
  filter(str_detect(
    subjects,
    regex("Ceasefires|Peacekeeping Operations|Internal Security|Arms Embargo", ignore_case = TRUE)
  ))

law_similarity <-
  vote_similarity(law)

# Human Rights

rights <- 
  sanctions_votes |> 
  filter(str_detect(
    subjects,
    regex("Human Rights|Human Rights in Armed Conflicts|Humanitarian Assistance|Displaced Persons", ignore_case = TRUE)
  ))

rights_similarity <-
  vote_similarity(rights)

```

#### Vote Similarity by Top 10 countries sanctioned

```{r Vote Similarity by Top 10 countries sanctioned}
#| code-fold: true

# Top 5 countries sanctioned: Iraq, Sudan, Côte d'Ivoire, Somalia, Libya

top_countries <- sanctions_votes |>
  separate_rows(names, sep = "\\|") |>
  filter(names != "") |>
  mutate(names = str_trim(names)) |>
  count(names, sort = TRUE) |>
  slice_max(n, n = 50)

top_countries


# Iraq: 67

iraq <- 
  sanctions_votes |>
  filter(str_detect(
    names,
    regex("Iraq", ignore_case = TRUE)
  ))

iraq_similarity <- 
  vote_similarity(iraq)


# Sudan: 47

sudan <- 
  sanctions_votes |>
  filter(str_detect(
    names,
    regex("Sudan", ignore_case = TRUE)
  ))

sudan_similarity <- 
  vote_similarity(sudan)


# Côte d'Ivoire: 38

cote <- 
  sanctions_votes |>
  filter(str_detect(
    names,
    regex("Côte d'Ivoire", ignore_case = TRUE)
  ))

cote_similarity <- 
  vote_similarity(cote)


# Somalia: 37

somalia <- 
  sanctions_votes |>
  filter(str_detect(
    names,
    regex("Somalia", ignore_case = TRUE)
  ))

somalia_similarity <- 
  vote_similarity(somalia)


# Libya: 35

libya <- 
  sanctions_votes |>
  filter(str_detect(
    names,
    regex("Libya", ignore_case = TRUE)
  ))

libya_similarity <- 
  vote_similarity(libya)


# Top 5 countries sanctioned: Iraq, Sudan, Côte d'Ivoire, Somalia, Libya

# Widest range of similarity is over Sudan (0.43)
# Narrowest range of similarity is over Côte d'Ivoire (unanimous)

```

## Visualizing Similarity

#### Vote Similarity Pentagon Function With Title

```{r Vote Similarity Pentagon Function with title}
#| code-fold: true

plot_similarity_t <- function(sim_tbl, title) {
  
edges <- sim_tbl |> 
  separate(pair, into = c("from", "to"), sep = "-") |> 
  mutate(weight = similarity)

eastern <- c("CHINA", "RUSSIA")
western <- c("US", "UK", "FRANCE")

vertices <- data.frame(
  name = unique(c(edges$from, edges$to)),
  stringsAsFactors = FALSE
) |> 
  mutate(p5 = ifelse(name %in% western, "western", "eastern"))

g <- graph_from_data_frame(edges, directed = FALSE, vertices = vertices)

edges <- edges |> 
  mutate(
    group = case_when(
      from %in% eastern & to %in% eastern ~ "east_east",
      from %in% western & to %in% western ~ "west_west",
      TRUE ~ "east_west"
    )
  )

# Pentagon layout

order <- c("CHINA", "RUSSIA", "UK", "FRANCE", "US")

n <- length(order)
radius <- 25
angles <- (90 + seq(0, 360 - 360/n, length.out = n)) * pi/180
coords <- tibble::tibble(
  name = order,
  x = cos(angles) * radius,
  y = sin(angles) * radius
)

layout <- create_layout(g, layout = "circle")
layout$x <- coords$x[match(layout$name, coords$name)]
layout$y <- coords$y[match(layout$name, coords$name)]

# Plotting

xlim <- c(min(layout$x) - 0.5, max(layout$x) + 0.5)
ylim <- c(min(layout$y) - 0.5, max(layout$y) + 0.5)

splot <- ggraph(layout, rescale = FALSE) +
  geom_edge_arc(
    aes(width = weight, colour = edges$group),
    alpha = 0.4,
    strength = 0.125
  ) +
  scale_edge_width(limits = c(0.6, 1), range = c(0, 12.5)) +
  guides(edge_width = "none", fill = "none", edge_colour = "none") +
  geom_node_point(
    aes(fill = p5),
    size = 35,
    alpha = 0.75,
    shape = 21,
    colour = "black",
    stroke = 0.5
  ) +
  scale_fill_manual(values = c(western = "steelblue", eastern = "indianred3")) +
  scale_edge_colour_manual(values = c(
    east_east = "red2",
    west_west = "royalblue",
    east_west = "mediumpurple2"
  )) +
  geom_node_text(
    aes(label = toupper(name)),
    size = 5.5,
    vjust = 0.5, hjust = 0.5,
    family = "serif",
    fontface = "bold"
  ) +
  theme_void() +
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    plot.margin = margin(20, 0, 40, 0),
    plot.title = element_text(
      face = "bold", hjust = 0.5, size = 16, margin = margin(b = 40,)
    )
  ) +
  coord_equal(xlim = xlim, ylim = ylim, clip = "off") +
  labs(title = title)

# Table of similarity values

  pair_labels <- c(
    "CHINA-US"     = "China-US",
    "US-CHINA"     = "China-US",
    "CHINA-FRANCE" = "China-France",
    "FRANCE-CHINA" = "China-France",
    "US-RUSSIA"    = "US-Russia",
    "RUSSIA-US"    = "US-Russia",
    "RUSSIA-FRANCE"= "Russia-France",
    "FRANCE-RUSSIA"= "Russia-France",
    "CHINA-RUSSIA" = "China-Russia",
    "RUSSIA-CHINA" = "China-Russia",
    "CHINA-UK"     = "China-UK",
    "UK-CHINA"     = "China-UK",
    "RUSSIA-UK"    = "Russia-UK",
    "UK-RUSSIA"    = "Russia-UK",
    "US-UK"        = "US-UK",
    "UK-US"        = "US-UK",
    "FRANCE-UK"    = "France-UK",
    "UK-FRANCE"    = "France-UK",
    "US-FRANCE"    = "US-France",
    "FRANCE-US"    = "US-France"
  )
  
  table_data <- sim_tbl |>
    mutate(
      Pair = pair_labels[pair],
      Similarity = round(similarity * 100, 2)
    ) |>
    filter(!is.na(Pair)) |>
    select(Pair, Similarity) |>
    arrange(Similarity)
  
  beige_light  <- "#F5F0E0"
  beige_dark   <- "#EDE6D7"
  header_beige <- "#D6CCB7"
  
  tbl <- 
    table_data |>
    gt() |>
    cols_label(
      Pair = "Pair",
      Similarity = "Similarity (%)"
    ) |>
    fmt_number(
      columns = Similarity,
      decimals = 2
    ) |>
    cols_align(align = "left",   columns = Pair) |>
    cols_align(align = "center", columns = Similarity) |>
    tab_style(
      style = cell_fill(color = "#F5F0E0"),
      locations = cells_body(rows = seq(1, nrow(table_data), 2))
    ) |>
    tab_style(
      style = cell_fill(color = "#EDE6D0"),
      locations = cells_body(rows = seq(2, nrow(table_data), 2))
    ) |>
    tab_style(
      style = list(
        cell_fill(color = "#D6CCB0"),
        cell_text(weight = "bold", font = "serif")
      ),
      locations = cells_column_labels()
    ) |>
    tab_style(
      style = cell_text(font = "serif"),
      locations = cells_body()
    ) |>
    tab_options(
      table.border.top.style = "none",
      table.border.bottom.style = "none",
      column_labels.border.bottom.width = px(1),
      column_labels.border.bottom.color = "#B0A898",
      table_body.hlines.style = "none",
      table.width = pct(100)
    )
  
  # Using with patchwork
  tbl_gg <- wrap_elements(full = tbl)
  
  combined <- splot + tbl_gg +
    plot_layout(widths = c(2, 1))
  
  print(combined)
}

```

#### Vote Similarity Pentagon Function Without Title

```{r Vote Similarity Pentagon Function without title}
#| code-fold: true

plot_similarity <- function(sim_tbl) {
  
edges <- sim_tbl |> separate(pair, into = c("from", "to"), sep = "-")

eastern <- c("CHINA", "RUSSIA")
western <- c("US", "UK", "FRANCE")

vertices <- data.frame(
  name = unique(c(edges$from, edges$to)),
  stringsAsFactors = FALSE
  ) |> 
  mutate(p5 = ifelse(name %in% western, "western", "eastern"))

g <- graph_from_data_frame(edges, directed = FALSE, vertices = vertices)

edges <- edges |> 
  mutate(
    group = case_when(
      from %in% eastern & to %in% eastern ~ "east_east",
      from %in% western & to %in% western ~ "west_west",
      TRUE ~ "east_west"
    )
  )

# Pentagon layout

order <- c("CHINA", "RUSSIA", "UK", "FRANCE", "US")

n <- length(order)
radius <- 20
angles <- (90 + seq(0, 360 - 360/n, length.out = n)) * pi/180
coords <- tibble::tibble(
  name = order,
  x = cos(angles) * radius,
  y = sin(angles) * radius
)

layout <- create_layout(g, layout = "circle")
layout$x <- coords$x[match(layout$name, coords$name)]
layout$y <- coords$y[match(layout$name, coords$name)]

# Plotting

xlim <- c(min(layout$x) - 0.5, max(layout$x) + 0.5)
ylim <- c(min(layout$y) - 0.5, max(layout$y) + 0.5)

splot <- ggraph(layout, rescale = FALSE) +
  geom_edge_arc(
    aes(width = similarity, colour = edges$group),
    alpha = 0.4,
    strength = 0.125
  ) +
  scale_edge_width(limits = c(0.6, 1), range = c(0, 12.5)) +
  guides(edge_width = "none", fill = "none", edge_colour = "none") +
  geom_node_point(
    aes(fill = p5),
    size = 35,
    alpha = 0.75,
    shape = 21,
    colour = "black",
    stroke = 0.5
  ) +
  scale_fill_manual(values = c(western = "steelblue", eastern = "indianred3")) +
  scale_edge_colour_manual(values = c(
    east_east = "red2",
    west_west = "royalblue",
    east_west = "mediumpurple2"
  )) +
  geom_node_text(
    aes(label = toupper(name)),
    size = 5.5,
    vjust = 0.5, hjust = 0.5,
    family = "serif",
    fontface = "bold"
  ) +
  theme_void() +
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    plot.margin = margin(40, 0, 40, 0),
    ) +
  coord_equal(xlim = xlim, ylim = ylim, clip = "off")

# Table of similarity values

  pair_labels <- c(
    "CHINA-US"     = "China-US",
    "US-CHINA"     = "China-US",
    "CHINA-FRANCE" = "China-France",
    "FRANCE-CHINA" = "China-France",
    "US-RUSSIA"    = "US-Russia",
    "RUSSIA-US"    = "US-Russia",
    "RUSSIA-FRANCE"= "Russia-France",
    "FRANCE-RUSSIA"= "Russia-France",
    "CHINA-RUSSIA" = "China-Russia",
    "RUSSIA-CHINA" = "China-Russia",
    "CHINA-UK"     = "China-UK",
    "UK-CHINA"     = "China-UK",
    "RUSSIA-UK"    = "Russia-UK",
    "UK-RUSSIA"    = "Russia-UK",
    "US-UK"        = "US-UK",
    "UK-US"        = "US-UK",
    "FRANCE-UK"    = "France-UK",
    "UK-FRANCE"    = "France-UK",
    "US-FRANCE"    = "US-France",
    "FRANCE-US"    = "US-France"
  )
  
  table_data <- sim_tbl |>
    mutate(
      Pair = pair_labels[pair],
      Similarity = round(similarity * 100, 2)
    ) |>
    filter(!is.na(Pair)) |>
    select(Pair, Similarity) |>
    arrange(Similarity)
  
  beige_light  <- "#F5F0E0"
  beige_dark   <- "#EDE6D7"
  header_beige <- "#D6CCB7"
  
  tbl <- 
    table_data |>
    gt() |>
    cols_label(
      Pair = "Pair",
      Similarity = "Similarity (%)"
    ) |>
    fmt_number(
      columns = Similarity,
      decimals = 2
    ) |>
    cols_align(align = "left",   columns = Pair) |>
    cols_align(align = "center", columns = Similarity) |>
    tab_style(
      style = cell_fill(color = "#F5F0E0"),
      locations = cells_body(rows = seq(1, nrow(table_data), 2))
    ) |>
    tab_style(
      style = cell_fill(color = "#EDE6D0"),
      locations = cells_body(rows = seq(2, nrow(table_data), 2))
    ) |>
    tab_style(
      style = list(
        cell_fill(color = "#D6CCB0"),
        cell_text(weight = "bold", font = "serif")
      ),
      locations = cells_column_labels()
    ) |>
    tab_style(
      style = cell_text(font = "serif"),
      locations = cells_body()
    ) |>
    tab_options(
      table.border.top.style = "none",
      table.border.bottom.style = "none",
      column_labels.border.bottom.width = px(1),
      column_labels.border.bottom.color = "#B0A898",
      table_body.hlines.style = "none",
      table.width = pct(100)
    )
  
  # Using with patchwork
  tbl_gg <- wrap_elements(full = tbl)
  
  combined <- splot + tbl_gg +
    plot_layout(widths = c(2, 1))
  
  print(combined)
}

```

#### Vote Similarity Chart Function With Title

```{r Vote Similarity Chart function with title}
#| code-fold: true

# Custom function to present similarity data with more detail

chart_similarity_t <- function(sim_tib, title) {
  
# Order by similarity

  sim_plot <- 
    sim_tib |> 
    mutate(
      pair = factor(pair, levels = pair[order(similarity)])
    )

# Plotting

  splot <- 
    ggplot(sim_plot, aes(x = pair, y = similarity)) +
    geom_col(fill = "steelblue", alpha = 0.9) +
    coord_flip() +
    scale_y_continuous(limits = c(0, 1.05), expand = c(0, 0)) +
    labs(
      title = title,
      x = NULL,
      y = "Vote Similarity Score"
    ) +
    theme_minimal(base_size = 14) +
    theme(
      plot.margin = margin(30, 15, 30, 15), #top, right, bottom, left
      plot.title = element_text(
        face = "bold", hjust = 0, size = 16, margin = margin(b = 20)
      ),
      axis.title.x = element_text(margin = margin(t = 10)),
      panel.grid.major.y = element_blank(),
      panel.grid.minor = element_blank(),
      panel.background = element_rect(fill = "white", color = NA),
      plot.background = element_rect(fill = "white", color = NA)
    )
  
  print(splot)
  
}

```

#### Vote Similarity Chart Function Without Title

```{r Vote Similarity Chart function without title}
#| code-fold: true

# Custom function to present similarity data with details

chart_similarity <- function(sim_tib) {
  
# Order by similarity

  sim_plot <- 
    sim_tib |> 
    mutate(
      pair = factor(pair, levels = pair[order(similarity)])
    )

# Plotting

  splot <- 
    ggplot(sim_plot, aes(x = pair, y = similarity)) +
    geom_col(fill = "steelblue", alpha = 0.9) +
    coord_flip() +
    scale_y_continuous(limits = c(0, 1.05), expand = c(0, 0)) +
    labs(
      x = NULL,
      y = "Vote Similarity Score"
    ) +
    theme_minimal(base_size = 14) +
    theme(
      plot.margin = margin(30, 15, 30, 15), #top, right, bottom, left
      axis.title.x = element_text(margin = margin(t = 10)),
      panel.grid.major.y = element_blank(),
      panel.grid.minor = element_blank(),
      panel.background = element_rect(fill = "white", color = NA),
      plot.background = element_rect(fill = "white", color = NA)
    )
  
  print(splot)
  
}
```

#### General Vote Similarity Plot

```{r General Vote Similarity plot}
#| code-fold: true

gen_sim_plot <- 
  plot_similarity(general_similarity)

gen_sim_chart <- 
  chart_similarity(general_similarity)

general_similarity

#ggsave(filename = "general_similarity_plot.png", plot = gen_sim_plot)
#ggsave(filename = "general_similarity_chart.png", plot = gen_sim_chart)
```

#### LIO Vote Similarity Plots

```{r LIO Vote Similarity Plots}
#| code-fold: true

law_sim_plot <- 
  plot_similarity(law_similarity)

law_sim_chart <- 
  chart_similarity(law_similarity)

law_similarity


rights_sim_plot <- 
  plot_similarity(rights_similarity)

rights_sim_chart <- 
  chart_similarity(rights_similarity)

rights_similarity


#ggsave(filename = "law_similarity_plot.png", plot = law_sim_plot)
#ggsave(filename = "law_similarity_chart.png", plot = law_sim_chart)
#ggsave(filename = "rights_similarity_plot.png", plot = rights_sim_plot)
#ggsave(filename = "rights_similarity_chart.png", plot = rights_sim_chart)

```

#### Vote Similarity Plots for Top 5 Countries

```{r Top 5 Vote Similarity Plot}
#| code-fold: true

# Iraq, Sudan, Côte d'Ivoire, Somalia, Libya

sudan_chart <- 
  chart_similarity(sudan_similarity)

somalia_chart <- 
  chart_similarity(somalia_similarity)

libya_chart <- 
  chart_similarity(libya_similarity)

iraq_chart <- 
  chart_similarity(iraq_similarity)

cote_chart <- 
  chart_similarity(cote_similarity)

#ggsave(filename = "sudan_chart.png", plot = sudan_chart)
#ggsave(filename = "somalia_chart.png", plot = somalia_chart)
#ggsave(filename = "libya_chart.png", plot = libya_chart)
#ggsave(filename = "iraq_chart.png", plot = iraq_chart)
#ggsave(filename = "cote_chart.png", plot = cote_chart)

```

## Syria Vetoes

```{r Syria}
#| code-fold: true

# 12 of China's 19 vetoes have been over the Middle East, 10 over Syria, 2 over Palestine

vetoed |> 
  filter(china_no == 1) |> count(topic, sort = TRUE)


# 10 Vetoes over Syria, 4 over sanctions

syria_veto <- 
  vetoed |>
  filter(china_no == 1) |> 
  filter(str_detect(
    topic, regex("syria", ignore_case = TRUE)
  ))

syria_veto

# 8 Resolutions Syria was involved (but not implicated) in, 1 sanctioned Syria (Res. 1636: Syria must detain bombing suspects, sanctions against implicated Syrian officials), with an unknown success rating

syria_votes <- 
  sanctions_votes |>
  filter(str_detect(
    names,
    regex("syria", ignore_case = TRUE)
  )) |> 
  slice(1)

syria_votes

# Out of 5 drafts to impose sanctions on Syria, one passed, imposing sanctions, while China vetoed the other 4, using 4 of the 19 vetoes it has ever cast. Due to China's vetoes, sanctioning Syria has a 1/5 chance of success, meaning China's support in sanctioning Syria is very important.

```

## Visualizing Vetoes

#### Veto Data

```{r Veto Data}
#| code-fold: true

# China's veto similarity with vetoing alongside, or against, the other states
veto_similarity <- 
  vetoed |>
  filter(china_no == 1) |> 
  mutate(
    us_d = case_when(us_no == 0 ~ 1, TRUE ~ 0),
    france_d = case_when(france_no == 0 ~ 1, TRUE ~ 0),
    russia_d = case_when(russia_no == 0 ~ 1, TRUE ~ 0),
    uk_d = case_when(uk_no == 0 ~ 1, TRUE ~ 0)
  ) |> 
  select(-year) |> 
  select(date, draft_no, topic, us_d, france_d, russia_d, uk_d, success)

veto_similarity

# Russia abstained on S/1999/201, so China's veto does not count against them. Therefore, between 1972-2024, China vetoed two resolutions that Russia voted "yes" on, and vetoed 19 resolutions that each of the other western p5 voted "yes" on. All of its vetoes were against the US, UK, and France.

# All 19 of China's vetoes were when all of the P5 were present as well (no nonvotes)


# Converting to long format

veto_sim_bar_data <- 
  veto_similarity |> 
  summarize(
    US = sum(us_d),
    France = sum(france_d),
    Russia = sum(russia_d) - 1,
    UK = sum(uk_d)
  ) |> 
  pivot_longer(
    cols = everything(),
    names_to = "State",
    values_to = "Count"
  ) |> 
  mutate(
    State = factor(State, levels = c("US", "Russia", "France", "UK"))
  )

veto_sim_bar_data

```

#### Veto Similarity Chart With Title

```{r Veto Similarity chart with title}
#| code-fold: true

veto_sim_graph_t <- 
  ggplot(veto_sim_bar_data, aes(x = State, y = Count, fill = State)) +
  geom_col(width = 0.7) +
  geom_text(aes(label = Count), vjust = -0.5, size = 5.5) +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 18),
    plot.margin = margin(15, 10, 10, 10),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA)
  ) +
  labs(
    title = "China's Vetoes Against Other P5 Members",
    subtitle = "China Vetoed 19 times from 1972-2024",
    x = NULL,
    y = "Count"
  ) +
  coord_cartesian(ylim = c(0, max(veto_sim_bar_data$Count) + 2))

veto_sim_graph_t


#ggsave(filename = "veto_sim.png", plot = veto_sim_graph)

```

#### Veto Similarity Chart Without Title

```{r Veto Similarity chart without title}
#| code-fold: true

veto_sim_graph <- 
  ggplot(veto_sim_bar_data, aes(x = State, y = Count, fill = State)) +
  geom_col(width = 0.7) +
  geom_text(aes(label = Count), vjust = -0.5, size = 5.5) +
  scale_fill_manual(values =
    c("US" = "#46549C", "Russia" = "#33a02c", "France" = "#0072BB", "UK" = "#C8102E")
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.margin = margin(15, 10, 10, 10),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA)
  ) +
  labs(
    x = NULL,
    y = "Count"
  ) +
  coord_cartesian(ylim = c(0, max(veto_sim_bar_data$Count) + 2))

veto_sim_graph

#ggsave(filename = "veto_sim.png", plot = veto_sim_graph)

```

## Timeline Visualtion

#### Timeline Data

```{r Timeline Data}
#| code-fold: true

# All cases with either a vote or abstain by China

china_vv_full <- 
  sanctions_votes |>
  filter(china_yes == 1 | china_no == 1 | china_ab == 1) |> 
  mutate(
    china_ns = if_else(china_no == 1 | china_ab == 1, 1, 0),
    china_yes = replace_na(china_yes, 0)
  ) |> 
  select(year, date, res_no, draft_no, topic, subjects, china_yes, china_ns)

china_vv_full

china_vv_full |> summarize(sum = sum(china_ns))
# China has withheld support from 65 sanctions, with 19 vetoes and 46 abstains
```

#### Timeline of China Support For Sanctions

```{r Timeline of China support for sanctions}
#| code-fold: true

china_yearly <- 
  china_vv_full |> 
  group_by(year) |> 
  summarize(
    support = sum(china_yes, na.rm = TRUE),
    nonsupport = sum(china_ns, na.rm = TRUE)
  ) |> 
  mutate(nonsupport = -nonsupport)   # non-support negative for diverging bars

china_yearly_graph <- 
  ggplot(china_yearly, aes(x = year)) +
  geom_col(aes(y = support), fill = "steelblue", width = 0.8) +
  geom_col(aes(y = nonsupport), fill = "firebrick", width = 0.8) +
  geom_hline(yintercept = 0, color = "black", linewidth = 0.6) +
  scale_x_continuous(
    breaks = seq(min(china_yearly$year), max(china_yearly$year), by = 5)
  ) +
  scale_y_continuous(
#    labels = abs,   # positive numbers on both sides
    expand = expansion(mult = c(0.05, 0.1))
  ) +
  labs(
    x = "Year",
    y = "Supported Sanctions",
    caption = "Upward bars = Support (Yes votes); Downward bars = Non-support (Abstain or Veto)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold"),
    axis.title.y = element_text(
      face = "bold",
      margin = margin(r = 10)
    ),
    axis.title.x = element_text(
      face = "bold",
      margin = margin(t = 15)
    ),
    plot.caption = element_text(
      margin = margin(t = 15)
    ),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA)
  )

china_yearly_graph

#ggsave(filename = "timeline.png", plot = china_yearly_graph)

```
